quollr: An R Package for Visualizing 2-D Models from Nonlinear Dimension Reductions in High-Dimensional Space

### 2025-12-14 paper-quollr.Rmd
Chose another binwidth

We illustrate how to use `quollr` to assess whether this is a reasonable layout. Figure \@ref(fig:limb-hbe) shows five alternative layouts, and the HBE plot summarizing the resulting model fits. Layout a is produce by UMAP (neighbors=5, minimum distance=0.1); layout c was produced by phate (knn=5); layout d was produced by trimap (number of inliers=12, outliers=4, random=3); layout e was produced by by pacmap (neighbors=10, init_random_MN-ratio=0.5, FP-ratio=2); layout f was produced by t-SNE (perplexity=15).

The HBE plot indicates that the two t-SNE layouts outperform all the other methods across a range of binwidths, but that the result with perplexity of 15 outperforms the other. There are small visual difference between the two layouts. Both support that there are 5-6 clusters. Layout b has slightly more space between clusters.  Layout d three small clusters at the top whereas layout f has only two, and another smaller one at the bottom. 
<!-- add function to generate 2D trimesh and the projections (PDF) and langevitour (HTML)-->

@@

tsne_limb_obj <- fit_highd_model(
  highd_data = data_limb, 
  nldr_data = tsne_limb, 
  b1 = 21, 
  q = 0.1, 
  hd_thresh = 0)

@@

`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:tsne-link-limb), \\@ref(fig:tsne-link-limb-best)"} else { "Figure \\@ref(fig:model-limb)"})` show how to examine the resulting models of the layouts overlaid on the data in high dimensions. Point color represents the cluster reported in the published paper. In each case the best model is produced using binwidth=XXX (and why). The plots in these figures are  best understand using small steps.

@@

tsne_best_limb_obj <- fit_highd_model(
  highd_data = data_limb, 
  nldr_data = tsne_limb2, 
  b1 = 21, 
  q = 0.1, 
  hd_thresh = 0)

>>>>

We illustrate how to use `quollr` to assess whether this is a reasonable layout. Figure \@ref(fig:limb-hbe) shows five alternative layouts, and the HBE plot summarizing the resulting model fits. Layout a is produce by UMAP (neighbors=5, minimum distance=0.1); layout c was produced by phate (knn=5); layout d was produced by trimap (number of inliers=12, outliers=4, random=3); layout e was produced by by pacmap (neighbors=10, init_random_MN-ratio=0.5, FP-ratio=2); layout f was produced by tSNE (perplexity=15).

The HBE plot indicates that the two tSNE layouts outperform all the other methods across a range of binwidths, but that the result with perplexity of 15 outperforms the other. There are small visual difference between the two layouts. Both support that there are 5-6 clusters. Layout b has slightly more space between clusters.  Layout d three small clusters at the top whereas layout f has only two, and another smaller one at the bottom. 
<!-- add function to generate 2D trimesh and the projections (PDF) and langevitour (HTML)-->

@@

tsne_limb_obj <- fit_highd_model(
  highd_data = data_limb, 
  nldr_data = tsne_limb, 
  b1 = 28, 
  q = 0.1, 
  hd_thresh = 0)

@@

`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:tsne-link-limb), \\@ref(fig:tsne-link-limb-best)"} else { "Figure \\@ref(fig:model-limb)"})` show how to examine the resulting models of the layouts overlaid on the data in high dimensions. Point color represents the cluster reported in the published paper. In each case the best model is produced using binwidth=0.06 (and why). The plots in these figures are  best understand using small steps.

@@

tsne_best_limb_obj <- fit_highd_model(
  highd_data = data_limb, 
  nldr_data = tsne_limb2, 
  b1 = 28, 
  q = 0.1, 
  hd_thresh = 0)

### 2025-12-14 paper-quollr.tex
Chose another binwidth

We illustrate how to use \texttt{quollr} to assess whether this is a reasonable layout. Figure \ref{fig:limb-hbe} shows five alternative layouts, and the HBE plot summarizing the resulting model fits. Layout a is produce by UMAP (neighbors=5, minimum distance=0.1); layout c was produced by phate (knn=5); layout d was produced by trimap (number of inliers=12, outliers=4, random=3); layout e was produced by by pacmap (neighbors=10, init\_random\_MN-ratio=0.5, FP-ratio=2); layout f was produced by t-SNE (perplexity=15).

The HBE plot indicates that the two t-SNE layouts outperform all the other methods across a range of binwidths, but that the result with perplexity of 15 outperforms the other. There are small visual difference between the two layouts. Both support that there are 5-6 clusters. Layout b has slightly more space between clusters. Layout d three small clusters at the top whereas layout f has only two, and another smaller one at the bottom.

@@

Figure \ref{fig:model-limb} show how to examine the resulting models of the layouts overlaid on the data in high dimensions. Point color represents the cluster reported in the published paper. In each case the best model is produced using binwidth=XXX (and why). The plots in these figures are best understand using small steps.

>>>>

We illustrate how to use \texttt{quollr} to assess whether this is a reasonable layout. Figure \ref{fig:limb-hbe} shows five alternative layouts, and the HBE plot summarizing the resulting model fits. Layout a is produce by UMAP (neighbors=5, minimum distance=0.1); layout c was produced by phate (knn=5); layout d was produced by trimap (number of inliers=12, outliers=4, random=3); layout e was produced by by pacmap (neighbors=10, init\_random\_MN-ratio=0.5, FP-ratio=2); layout f was produced by tSNE (perplexity=15).

The HBE plot indicates that the two tSNE layouts outperform all the other methods across a range of binwidths, but that the result with perplexity of 15 outperforms the other. There are small visual difference between the two layouts. Both support that there are 5-6 clusters. Layout b has slightly more space between clusters. Layout d three small clusters at the top whereas layout f has only two, and another smaller one at the bottom.

@@

Figure \ref{fig:model-limb} show how to examine the resulting models of the layouts overlaid on the data in high dimensions. Point color represents the cluster reported in the published paper. In each case the best model is produced using binwidth=0.06 (and why). The plots in these figures are best understand using small steps.
